---
title: "Introduction to Plotting"
author: "R Team"
output:
  pdf_document: default
---

# Data Visualization 
- Widely used concept in business
- Communication between analysts and stakeholders
- We will be using ggplot2 for graphing, other options are possible in R
- For further readings: http://www.cookbook-r.com/Graphs/
- We will cover different type of charts:
  1. Scatter Plot
  2. Line Charts
  3. Box plots
  4. Histograms
  
**Note**: Use View(data_frame) to choose which columns to use as x and y.

```{r message=FALSE, echo=TRUE}
# Importing libraries
library(palmerpenguins)
library(ggplot2)
library(tidyverse) 

# Importing datasets
spotify_songs <- read.csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
movie_profit <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-10-23/movie_profit.csv")
housing_df <- read_csv("https://www.dropbox.com/s/tvvtf9dwjufo7os/housing_train.csv?dl=1") 
penguins <- palmerpenguins::penguins
```

# 1. Scatter Plot 

**ggplot(data_frame, aes(x,y)) + geom_point()**:

  - First argument is the dataframe we are graphing
  - Second argument identifies the x and y coordinates using the aes function
  - Then different type of properties are adding by adding to the initial definition

```{r out.width="200px", out.height="200px"}
ggplot(housing_df, aes(x=MSSubClass, y=LotArea)) + geom_point()
```

### Can you see a relation between energy and track_popularity?
- Plots can be helpful in identifying certain correlations.
- One could visualize data and extract insights just by looking at the graphs.

```{r out.width="200px", out.height="200px"}
ggplot(spotify_songs, aes(x=energy, y=track_popularity)) + geom_point()
```

**Exercise**: 

1. Scatter plot of domestic_gross (x) vs worldwide_gross (y)
2. Scatter plot of MasVnrArea (x) vs LotArea (y)

# 2. Line Chart 
- When the lines of the scatter plot are connected

```{r out.width="200px", out.height="200px"}
ggplot(housing_df, aes(x=YearBuilt, y=SalePrice)) + geom_line()
```
  
  * The geom_line() is the part which identifies the chart as a line chart.
  * The rest of the function is the same, we specify the dataframe and x,y columns.
  * Plotting the data YearBuilt as x and SalePrice as y.

**Exercise**: 

1. Scatter plot of domestic_gross (x) vs worldwide_gross (y)
2. Scatter plot of MasVnrArea (x) vs LotArea (y)    
    
# 3. Bar Chart 
- It has the same structure but instead of geom_point we use geom_bar() with its identity parameter set to identity
    
```{r out.width="200px", out.height="200px"}
ggplot(BOD, aes(x=Time, y=demand)) + geom_bar(stat='identity')
```
    
- **Counting repetition**: We are interested to know which movie genre has the most movies (which genre has the highest repetition):
  1. First we pass movie_profit as the dataframe
  2. We choose genre as our x since we are interested in knowing the repetition
  3. We add geom_bar() to identify the graph as a bar chart.
        
```{r out.width="200px", out.height="200px"}
ggplot(movie_profit, aes(x = genre )) + geom_bar()
```        
    
- Imagine that we are interested in knowin what are the repetitions of the rating types within each genre. We could do this using an attribute called fill:
  1. We have followed the same process for steps 1 to 3.
  2. ***What does fill do?*** It groups data based on rating. We have already grouped data into genres, fill will group each genre group based on its ratings.
      
```{r out.width="200px", out.height="200px"}
ggplot(movie_profit, aes(x = genre, fill = mpaa_rating )) + geom_bar()
```        


**Note**: The position value with geom_bar() should either be dodge or fill

**Exercise**: (ET: 3/5)

  1. Plot clarity column of diamonds dataframe with cut as the fill.
  2. Plot clarity column of diamonds dataframe with color as the fill.


# 4. Histograms 
- Bar charts with connected bars

```{r out.width="200px", out.height="200px"}
ggplot(movie_profit, aes(x=log(worldwide_gross) )) + geom_histogram(binwidth = 2)
```      
  
  1. dataframe: we specify which dataframe we are interested in
  2. aes(x=log(worldwide_gross)): mapping the needed x and y values
  3. + geom_histogram(): Generates a histogram 
      
